Differences in the Risk of Grade Retention for Biracial and Monoracial Students in the US, 2010-2019

### 2023-03-14 products/paper/main.qmd
Rework conclusions to address issue #36

Differences in family resources were expected to play some role in the relative position of biracial students' risk. The results on parent selection vary somewhat by group, but the majority of biracial groups had resources more similar to their monoracial constituent group with higher resources, indicating a strong positive selection into interracial unions in the prior generation. The most important exception to this trend is the case of Black/White biracial students who have far fewer resources than white students and resources closer to black students.

@@

Because these findings rely upon large samples, they present an important step forward in our understanding of how biracial individuals will fit into America's system of racial inequality. However, this study also has important limitations that need to be noted.

Sample sizes for some non-white biracial respondents remain small, limiting our understanding of the outcomes for those groups. In particular, results for non-white part-Indigenous populations are quite imprecise. The results interestingly and counterintuitively suggest that part-Indigenous populations are the least likely to be in the position expected by hyperdescent, but due to low sample sizes, that finding is highly tentative.

Wile grade retention is an important early life outcome, it is far from the only one that these individuals will face in their life. Although, we often observe similar racial inequalities across a variety of outcomes, knowing the result for one outcome does not perfectly predict other outcomes. As biracial individuals age and face later outcomes in the educational system, the labor market, and elsewhere, their experiences may differ from what I observe here. Thus, collecting more detailed information on these outcomes in future studies remains critically important.

>>>>

Differences in family resources were expected to play some role in the relative position of biracial students' risk. The results on parent selection vary somewhat by group, but the majority of biracial groups had resources more similar to their monoracial constituent group with higher resources, indicating a strong positive selection into interracial unions in the prior generation. The most important exception to this trend is the case of Black/White biracial students who have far fewer resources than White students and resources closer to Black students.

@@

These findings stand in contrast to earlier work which found more evidence of an "in-between" status in other educational outcomes [@herman_blackwhiteother_2009; @campbell_multiracial_2009]. I raise two possible explanations for this discrepancy that point to strengths and weaknesses of the current research and may help drive future work on the topic. First, this discrepancy may be due to the small sample sizes of multiracial respondents in prior work which led to statistically imprecise estimates. Because the current findings rely upon large samples, they present an important step forward in our understanding of how biracial individuals will fit into America's system of racial inequality. Nonetheless, sample sizes for some non-white biracial respondents remain small, limiting our understanding of the outcomes for those groups. In particular, results for non-white part-Indigenous populations are quite imprecise. The results counter-intuitively suggest that part-Indigenous populations are the least likely to be in the position expected by hyperdescent, but due to low sample sizes, that finding is highly tentative.

Second, The current findings may also diverge from the findings of prior work because different outcomes were examined. While grade retention is an important early life outcome, it is far from the only one that these individuals will face in their life. Although, we often observe similar racial inequalities across a variety of outcomes, knowing the result for one outcome does not perfectly predict other outcomes. As biracial individuals age and face later outcomes in the educational system, the labor market, and elsewhere, their experiences may differ from what I observe here. Thus, collecting more detailed information on these outcomes in future studies remains critically important.
